Supplementary Material for \emph{A Guide to Pre-processing High-Frequency Animal Tracking Data}

### 2021-02-15 docs/ms_atlas_preproc_main_text.tex
update tex

PRG would like to thank Pedro M. Santos Neves for introducing PRG to \texttt{R} package development, for help with setting up \texttt{atlastools}, and for help with archiving it on Zenodo; 
Aparajitha Ramesh and Franjo Weissing for discussions on error propagation;
Geert Aarts, Jacob RL Gismann, Evy Gobbens for feedback that improved the manuscript; 
members of the Modelling Adaptive Response Mechanisms Group (Weissing Lab), and the Theoretical Biology department at the University of Groningen for helpful discussions on \texttt{atlastools} and the manuscript.
All authors thank the attendees of ATLAS workshops held in May and June 2020 at the Hebrew University of Jerusalem for helpful comments on the pipeline and \texttt{atlastools}.

>>>>

PRG would like to thank Pedro M. Santos Neves for introducing PRG to \texttt{R} package development, for help with setting up \texttt{atlastools}, and for help with archiving it on Zenodo; 
Aparajitha Ramesh and Franjo Weissing for discussions on error propagation;
Geert Aarts, Jacob RL Gismann, Evy Gobbens, and Roos Kentie for feedback that improved the manuscript; 
members of the Modelling Adaptive Response Mechanisms Group (Weissing Lab), and the Theoretical Biology department at the University of Groningen for helpful discussions on \texttt{atlastools} and the manuscript.
We thank the many volunteers, students, and NIOZ staff involved in operating the WATLAS tracking system, and most importantly Frank van Maarseveen, Bas Denissen and Anne Dekinga.
We also thank the attendees of ATLAS workshops held in May and June 2020 at the Hebrew University of Jerusalem for helpful comments on the pipeline and \texttt{atlastools}.
This work was partly funded by the Dutch Research Council grant VI.Veni.192.051 awarded to AIB.

### 2021-02-15 docs/ms_atlas_preproc_supplement.tex
update tex

\title{Supplementary Material for \emph{A Guide to Pre-processing High-Frequency Animal Tracking Data}}
\author{Pratik R. Gupte \and Christine E. Beardsworth \and Orr Spiegel \and Emmanuel Lourie \and Sivan Toledo \and Ran Nathan \and Allert I. Bijleveld}
\date{2021-02-11}

@@

Here we show how the residence patch method (Barraquand and Benhamou \protect\hyperlink{ref-barraquand2008}{2008}; Bijleveld et al. \protect\hyperlink{ref-bijleveld2016}{2016}; Oudman et al. \protect\hyperlink{ref-oudman2018}{2018}) accurately estimates the duration of known stops in a track collected as part of a calibration exercise in the Wadden Sea.

@@

@@

First we access the data from a local file using the \texttt{data.table} package (Dowle and Srinivasan \protect\hyperlink{ref-dowle2020}{2020}).
We then visualise the raw data.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# read and plot example data}
\NormalTok{data <-}\StringTok{ }\KeywordTok{fread}\NormalTok{(}\StringTok{"data/atlas1060_allTrials_annotated.csv"}\NormalTok{)}
\NormalTok{data_raw <-}\StringTok{ }\KeywordTok{copy}\NormalTok{(data)}
\end{Highlighting}
\end{Shaded}

@@

An indicator of individual residence at or near a position can be useful when attempting to identify residence patches. Positions can be filtered on a metric such as residence time (Bracis, Bildstein, and Mueller \protect\hyperlink{ref-bracis2018}{2018}).

@@

@@

@@

\hypertarget{plot-figure-7}{%
\section{Plot figure 7}\label{plot-figure-7}}

@@

>>>>

\title{Supplementary Material for \emph{A Guide to Pre-processing High-Frequency Animal Tracking Data}}
\author{Pratik R. Gupte \and Christine E. Beardsworth \and Orr Spiegel \and Emmanuel Lourie \and Sivan Toledo \and Ran Nathan \and Allert I. Bijleveld}
\date{2021-02-15}

@@

Here we show how the residence patch method (Barraquand and Benhamou \protect\hyperlink{ref-barraquand2008}{2008}; Bijleveld et al. \protect\hyperlink{ref-bijleveld2016}{2016}; Oudman et al. \protect\hyperlink{ref-oudman2018}{2018}) accurately estimates the duration of known stops in a track collected as part of a calibration exercise in the Wadden Sea. These data can be accessed from the \texttt{data} folder at this link: \url{https://doi.org/10.5281/zenodo.4287462}. These data are more fully reported in (Beardsworth et al. \protect\hyperlink{ref-beardsworth2021}{2021}).

@@

\hypertarget{a-note-on}{%
\subsection*{\texorpdfstring{A Note on \texttt{:=}}{A Note on :=}}\label{a-note-on}}
\addcontentsline{toc}{subsection}{A Note on \texttt{:=}}

The \texttt{atlastools} package is based on \texttt{data.table}, to be fast and efficient (Dowle and Srinivasan \protect\hyperlink{ref-dowle2020}{2020}).
A key feature is modification in place, where data is changed without making a copy.
This is already implemented in \texttt{R} and will be familiar to many users as \texttt{data\_frame\$column\_name\ \textless{}-\ values}.

The \texttt{data.table} way of writing this assignment would be \texttt{data\_frame{[},\ column\_name\ :=\ values{]}}.
We use this syntax throughout, as it provides many useful shortcuts, such as multiple assignment:

\texttt{data\_frame{[},\ c("col\_a",\ "col\_b)\ :=\ list(values\_a,\ values\_b){]}}

Users can use this special syntax, and will find it convenient with practice, but there are \emph{no} cases where users \emph{must} use the \texttt{data.table} syntax, and can simply treat the data as a regular \texttt{data.frame}. However, users are advised to convert their \texttt{data.frame} to a \texttt{data.table} using the function \texttt{data.table::setDT()}.

@@

First we access the data from a local file using the \texttt{data.table} package (Dowle and Srinivasan \protect\hyperlink{ref-dowle2020}{2020}).
We look at the first few rows, using \texttt{head()}.
We then visualise the raw data.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# read and plot example data}
\NormalTok{data <-}\StringTok{ }\KeywordTok{fread}\NormalTok{(}\StringTok{"data/atlas1060_allTrials_annotated.csv"}\NormalTok{)}
\NormalTok{data_raw <-}\StringTok{ }\KeywordTok{copy}\NormalTok{(data)}

\CommentTok{# see raw data}
\KeywordTok{head}\NormalTok{(data_raw)}
\end{Highlighting}
\end{Shaded}

@@

First, we filter data where we know the animal spent some time at or near a position, as this is the first step to identify residence patches.
One way of doing this is to filter out positions with speeds above which the animal is likely to be in transit, but a better way of doing this by filtering on a metric such as residence time (Bracis, Bildstein, and Mueller \protect\hyperlink{ref-bracis2018}{2018}).

@@

\hypertarget{a-note-on-summary-statistics}{%
\subsection*{A note on summary statistics}\label{a-note-on-summary-statistics}}
\addcontentsline{toc}{subsection}{A note on summary statistics}

Users specifying a \texttt{summary\_variable} should make sure that the variable for which they want a summary statistic is present in the data.
For instance, requesting mean speed by passing \texttt{summary\_variable\ =\ "speed"} and \texttt{summary\_function\ =\ "mean"} to \texttt{atl\_res\_patch}, should make sure that their data includes a column called \texttt{speed}.

@@

At this stage, users have successfully pre-processed their data from raw positions to residence patches.
Residence patches are essentially \texttt{sf} objects and can be visualised using the \texttt{sf} method for \texttt{plot}; for instance \texttt{plot(patch\_sf\_data)}.
Further sections reproduce the analyses in the main manuscript.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

@@

\hypertarget{main-text-figure-6}{%
\section{Main text Figure 6}\label{main-text-figure-6}}

@@

\leavevmode\hypertarget{ref-beardsworth2021}{}%
Beardsworth, Christine E., Evy Gobbens, Frank van Maarseveen, Bas Denissen, Anne Dekinga, Ran Nathan, Sivan Toledo, and Allert I. Bijleveld. 2021. ``Validating a High-Throughput Tracking System: ATLAS as a Regional-Scale Alternative to GPS.'' \emph{bioRxiv}, February, 2021.02.09.430514. \url{https://doi.org/10.1101/2021.02.09.430514}.
